Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-02-11 paper/paper.tex
More updates to paper.

\begin{figure}
\begin{verbatim}

@@

\end{verbatim}
\caption{A Persistent Linked List of Integers}
\label{fig:example}
\end{figure}

@@

Cyclone provides tools for programmers to write code that is failure
atomic. Portions of the code are delimited as transactions that are either
completely executed in terms of its effects on persistent memory or not executed
at all. Under the hood, Cyclone tracks all updates to persistent memory in a
persistent undo log - the log entries are persisted before the update to the
actual memory is allowed. On a failure, the undo log is used to roll back
incomplete transactions. A transaction is considered complete when it completes
execution and the updated objects are flushed (automatically by Cyclone) to
persistent memory. The log entries for the transaction are then immediately
deleted. The only burden placed on a programmer for failure atomicity is that
they must inform the Cyclone runtime about objects that are
updated. Figure~\ref{fig:example_fatomic} shows how to convert the function for
adding a node to a linked list into a failure atomic one. The only additions are
the transaction delimiters and the {\tt TX\_ADD} calls to mark the objects being
updated \emph{before} the actual updates to them.

\begin{figure}
\begin{verbatim}

@@

\end{verbatim}
\caption{A Persistent Linked List of Integers}
\label{fig:example}
\end{figure}

@@

>>>>

\begin{figure}
{\scriptsize
\begin{verbatim}

@@

\end{verbatim}
}
\caption{A Persistent Linked List of Integers}
\label{fig:example}
\end{figure}

@@

Cyclone provides tools for programmers to write code that is failure
atomic. Portions of the code are delimited as transactions that are either
completely executed in terms of its effects on persistent memory or not executed
at all. Under the hood, Cyclone tracks all updates to persistent memory in a
persistent undo log - the log entries are persisted before the update to the
actual memory is allowed. On a failure, the undo log is used to roll back
incomplete transactions. A transaction is considered complete when it completes
execution and the updated objects are flushed (automatically by Cyclone) to
persistent memory. The log entries for the transaction are then immediately
deleted. The only burden placed on a programmer for failure atomicity is that
they must inform the Cyclone runtime about objects that are
updated. Figure~\ref{fig:example_fatomic} shows how to convert the function for
adding a node to a linked list into a failure atomic one. The only additions are
the transaction delimiters and the {\tt TX\_ADD} calls to mark the objects being
updated \emph{before} the actual updates to them. Figure~\ref{fig:undo_log}
shows how Cyclone maintains the undo-log of objects.

\begin{figure}
{\scriptsize
\begin{verbatim}

@@

\end{verbatim}
}
\caption{A Persistent Linked List of Integers}
\label{fig:example_fatomic}
\end{figure}

\begin{figure}
{\scriptsize
\begin{verbatim}
TX_BEGIN()
{
  Initialize undo log
}

TX_ADD(o)
{
  if(o not already in undo log) {
    Append <o, value of o> to undo log
  }
}

TX_END()
{
  for each entry <o, v> in undo log {
    flush o to persistent storage
  }
  truncate undo log
}

Recover()
{
  for each entry <o, v> in undo log {
    set o to v
    flush o to persistent storage
  }   
}

\end{verbatim}
}
\caption{Undo Log in Cyclone}
\label{fig:undo_log}
\end{figure}

A crucial feature of Cyclone is that it abstracts the details of the persistent
mechanism away from programmers. This allows the same code to be deployed on
clusters with a persistent storage medium such as 3D XPoint or on clusters with
battery backed NVDIMMs for example. In the former case, explicit {\tt PCOMMIT}
instructions are necessary for persistence [ref PMFS paper], while in the latter
case no such instructions are necessary as on failure the CPU caches and DRAM
are flushed to stable storage under battery (or capacitor) power. Even in the
NVDIMM case however, the undo log is necessary to protect against OS or
application crashes to ensure that the persistent memory is always left in a
consistent state.

@@

\section{Related Work}
